---
title: "CFI_analysis"
output: pdf_document
date: "2025-02-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(tibble)
library(dplyr)
library(tidyr)
library(vegan)

getwd()

```

If needed: 

#getwd()
#setwd("/cloud/project/QB_biodiversity_project_EH/")

rm(list = ls())

#tree data 
tree <-read.csv("TREE.csv")

tree.species.df <-data.frame(Plot_ID = c(tree$PLOT),
                             SPCD =c(tree$SPCD))
str(tree.species.df) #plot and species id 
#print(tree.species.df)
#print(unique(tree.species.df$SPCD))

#Myco assocaitions
myco <-as.data.frame(read_xlsx("MycoType_ref2.xlsx"))
myco<- myco[,-c(3:7)]
#str(myco)
#print(unique(myco$SPCD))


#invasive
invasive <-read_xlsx("INVASIVE.xlsx")
#str(invasive)
invasive <-invasive
#print(invasive)
invasive.df <-as.data.frame(invasive$PLOT)  
#print(unique(invasive.df))

invasive.species.df <-data.frame(Plot_ID = c(invasive$PLOT),
                             SPCD =c(invasive$SP_ID))
str(invasive.species.df)

#References
#Plot ID Code 
plotID.master <-read_xlsx(path = "PLOT.xlsx")
#print(plotID.master)

#Species ID
speciesID.master <-read_xlsx(path = "REF_SPECIES.xlsx")
#print(speciesID.mater)

#Clean- make sure has same species
untree.species <- unique(tree.species.df$SPCD)
unmyco.species <-unique(myco$SPCD)
length(untree.species)
length(unmyco.species)
#print(sort(untree.species))
#print(sort(unmyco.species))
setdiff(untree.species, unmyco.species)

tree.species.df <- tree.species.df %>%
  filter(!SPCD %in% c(999, 998))  


#Filter if species has invasives 
#print(invasive.df)
invasive_plots <- unique(invasive.df$`invasive$PLOT`)
#print(invasive_plots)


invasive.plots.df <- data.frame()
ninvasive.plots.df <- data.frame()

for (i in seq_len(nrow(tree.species.df))) {  
  if (tree.species.df$Plot[i] %in% invasive_plots) {  
    invasive.plots.df <- rbind(invasive.plots.df, tree.species.df[i, ])  
  } else {  
    ninvasive.plots.df <- rbind(ninvasive.plots.df, tree.species.df[i, ])  
  }
}
# print(invasive.plots.df)
# print(ninvasive.plots.df)

#AM/ECM

ninvasive.myco.df <- merge(ninvasive.plots.df, myco, by="SPCD", all.x=TRUE, all.y=FALSE)

ninvasive.ECM.df <- subset(ninvasive.myco.df, ninvasive.myco.df$MycoType==1)
ninvasive.AM.df <- subset(ninvasive.myco.df, ninvasive.myco.df$MycoType==2)
ninvasive.both.df <- subset(ninvasive.myco.df, ninvasive.myco.df$MycoType==3)

invasive.myco.df <-merge(invasive.plots.df, myco, by="SPCD", all.x=TRUE, all.y=FALSE)
invasive.ECM.df <- subset(invasive.myco.df, invasive.myco.df$MycoType==1)
invasive.AM.df <- subset(invasive.myco.df, invasive.myco.df$MycoType==2)
invasive.both.df <- subset(invasive.myco.df, invasive.myco.df$MycoType==3)

#Site by species noninvasive

ninvasive.AM.ss <- as.data.frame.matrix(table(ninvasive.AM.df$Plot_ID, ninvasive.AM.df$SPCD))
ninvasive.AM.ss <- rownames_to_column(ninvasive.AM.ss, var = "Plot_ID")

ninvasive.ECM.ss <- as.data.frame.matrix(table(ninvasive.ECM.df$Plot_ID, ninvasive.ECM.df$SPCD))
ninvasive.ECM.ss <- rownames_to_column(ninvasive.ECM.ss, var = "Plot_ID")

ninvasive.both.ss <- as.data.frame.matrix(table(ninvasive.both.df$Plot_ID, ninvasive.both.df$SPCD))
ninvasive.both.ss <- rownames_to_column(ninvasive.both.ss, var = "Plot_ID")

#Site by species: Invasive
invasive.AM.ss <- as.data.frame.matrix(table(invasive.AM.df$Plot_ID, invasive.AM.df$SPCD))
invasive.AM.ss <- rownames_to_column(invasive.AM.ss, var = "Plot_ID")

invasive.ECM.ss <- as.data.frame.matrix(table(invasive.ECM.df$Plot_ID, invasive.ECM.df$SPCD))
invasive.ECM.ss <- rownames_to_column(invasive.ECM.ss, var = "Plot_ID")

invasive.both.ss <- as.data.frame.matrix(table(invasive.both.df$Plot_ID, invasive.both.df$SPCD))
invasive.both.ss <- rownames_to_column(invasive.both.ss, var = "Plot_ID")


#OVERALL 
ninvasive.AM.ss
ninvasive.ECM.ss
ninvasive.both.ss

invasive.AM.ss
invasive.ECM.ss
invasive.both.ss



```{r}
# Load data
PLOT <- read_excel("PLOT.xlsx")
SPCD <-PLOT$SPCD


# Prepare data frames with counts of distinct species per plot
tree_species_df <- tree %>%
  select(PLOT, SPCD) %>%
  group_by(PLOT) %>%
  summarise(Tree_Count = n_distinct(SPCD))

invasive_species_df <- invasive %>%
  select(PLOT, SPCD = SP_ID) %>%
  group_by(PLOT) %>%
  summarise(Invasive_Count = n_distinct(SPCD))

# Ensure all plots are represented
all_tree_df <- all_plots %>%
  left_join(tree_species_df, by = "PLOT") %>%
  replace_na(list(Tree_Count = 0))

all_invasive_df <- all_plots %>%
  left_join(invasive_species_df, by = "PLOT") %>%
  replace_na(list(Invasive_Count = 0))

# Assuming 'all_plots' contains all plot IDs from the "PLOT.xlsx" file
all_plots <- read_excel("PLOT.xlsx", col_names = TRUE)

# Create full matrices with zeros where data is missing
tree_full <- all_plots %>%
  left_join(tree_matrix, by = "PLOT") %>%
  replace_na(list(n = 0)) %>%
  select(-PLOT) %>%
  as.matrix()

invasive_full <- all_plots %>%
  left_join(invasive_matrix, by = "PLOT") %>%
  replace_na(list(n = 0)) %>%
  select(-PLOT) %>%
  as.matrix()



# Perform db-RDA
result <- dbrda(tree_matrix ~ invasive_matrix)


```



